ANGSD-wrapper: utilities for analyzing next generation sequencing data

### 2016-01-29 manuscript.tex
Minor changes to citations

A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}. 
However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
Korneliussen et al. \ph{Should be just \(2014\)} \citep{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}. 
ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage.

Here we present ANGSD-wrapper, a user-friendly interface to stable versions of ANGSD. 
ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.  
We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.

@@

The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope
This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
Each tab in the GUI contains plots for different ANGSD methods.

>>>>

A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}. 
However, few of these tools offer the flexibility to handle data with the characteristics of high throughput sequence data. 
\citet{korneliussen2014angsd} recently published the software package, ANGSD, which enables users to flexibly perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}. 
ANGSD works directly with the alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a niche format. 
One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than only on pre-determined variable sites. 
This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage.

Here we present ANGSD-wrapper, a user-friendly interface to stable versions of ANGSD. 
ANGSD-wrapper takes the form of a set of configuration files and `wrapper' scripts (Figure \ref{fig:supp2}) that streamline the execution of multi-step pipelines required for data analysis in ANGSD. 
ANGSD-wrapper also eases the configuration of ANGSD-related programs such as ngsPopGen \citep{ngsPopGen}, ngsF\citep{vieira2013estimating}, and ngsAdmix \citep{pmid24026093}. 
\ph{This could be reworded} Furthermore, ANGSD-wrapper allows for easy population genetic analysis across the sequence data provided, or hone in on a certain region or subset of regions, allowing a highly-tailored experience.
Additionally, the wrapper scripts are written against ``frozen'' versions of ANGSD and supporting tools, for consistency of analysis. 
Because the large volume of data associated with high throughput sequence analysis is often difficult to visualize, ANGSD-wrapper also provides a suite of interactive visualization tools to plot results and explore patterns at multiple scales.  
We demonstrate some of the analyses possible using ANGSD-wrapper when applied to low-coverage whole-genome data from domesticated maize and two related wild teosinte subspecies. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.

@@

\ph{Necessary?}The visualization software included with ANGSD-wrapper is contained within it's own directory called ``shinyGraphing.''
This application is started from within ANGSD-wrapper and launched locally from a standard web browser. %This application must be started in R and can be accessed locally from a web browser. --Nope
This software provides a graphical user interface (GUI) to quickly and interactively plot results obtained from ANGSD-wrapper.  
Each tab in the GUI contains plots for different ANGSD methods.
